### 2014-05-25 Supplement.tex
edits
edits. having trouble with supplement, trying to figure out issues

%Online Supplemental Materials
%Authors may wish to augment their manuscripts with online supplemental materials (e.g., large data sets, three-dimensional reconstructions, simulations, real-time movies, color photographs). Upload these appendices as separate files with the initial manuscript submission. Include a header on each file using this format: Smith et al.?American Journal of Botany 99(#): ###-###. 2012. ? Data Supplement S1 ? Page 1?. Name online supplements Appendix S1, Appendix S2, etc, in the order in which they appear in the text, regardless of whether they are tables, figures, text, other media, or a combination thereof. In the manuscript, after the mention of an online appendix, include the following: ?(see Supplemental Data with the online version of this article)?.

%Note that if authors wish to submit long DNA sequence appendices as supplemental material, they should select the "DNA sequences (online-only supplemental)" option on Editorial Manager. This ensures that lengthy appendices are not built into the reviewers' PDF, but are still accessible to the reviewers.
\documentclass[11pt]{article}
\usepackage{geometry} 
\usepackage{hyperref}
\usepackage[leftcaption]{sidecap}
\sidecaptionvpos{figure}{c} %position side caption
\usepackage{lineno}
\usepackage{lipsum}
\usepackage{indentfirst}
\usepackage{color}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage[pdftex]{graphicx}  
\usepackage{caption}
\usepackage{natbib}
\usepackage{authblk}
\usepackage[table]{xcolor} %for table line color alteration
\usepackage{subscript}
\begin{document}

\flushleft{\large{Vann \emph{et al.}: Variation in teosinte at the \emph{tb1} locus}}

% Figure S1. Cartoon representation of tb1 locus
%-------------------------------------------------------------------
\begin{figure*}[!t]
  \begin{center}
   \includegraphics[width=140mm]{FigS1LocusCartoon.pdf}
   \caption{Representation of the upstream regulatory region of \emph{tb1}, showing the \emph{tb1} coding region (green) and the \emph{Hopscotch} insertion (red). Arrows show the location of primer sets; in black, primers used for amplification and sequencing (Sequenced region 1; within the 5' UTR, and sequenced region 2; 66,169 bp upstream from the tb1 ORF); in blue, primers used to genotype the \emph{Hopscotch} insertion.}
\label{FigS1Locus}
  \end{center}
\end{figure*}
%-------------------------------------------------------------------

% Figure S2. Gel image
%-------------------------------------------------------------------
\begin{figure*}[!t]
  \begin{center}
   \includegraphics[width=140mm]{FigS2Gel.png}
   \caption{Agarose gel image of amplification products using our primer sets. Genotypes are indicated at the top of the gel.} 
\label{FigS2Gel}
  \end{center}
\end{figure*}
%-------------------------------------------------------------------

% Figure S3. NJ Trees
%-------------------------------------------------------------------
\begin{figure*}[!t]
  \begin{center}
   \includegraphics[width=140mm]{FigS3NJtrees.pdf}
    \caption{Neighbor-Joining tree of the sequenced region in the 5? UTR (right; sequenced region 1) and the 66,169 bp upstream region (left; sequenced region 2) of \emph{tb1}. Individuals with genotype data are colored: Homozygous for the teosinte (no \emph{Hopscotch}) allele (red), homozygous for the maize (\emph{Hopscotch}) allele (blue), heterozygotes (purple). TILs (teosinte inbred lines) are colored in green, with stars indicating the 3 TILs known to have the \emph{Hopscotch} insertion. Black indicates individuals not genotyped for the \emph{Hopscotch} insertion.} 
\label{FigS3NJ}
  \end{center}
\end{figure*}
%-------------------------------------------------------------------

% Table S1
\begin{table}[htbp]
\rowcolors{2}{gray!25}{white}
  \centering
  \caption{Pairwise F\textsubscript{ST} values from sequence and \emph{Hopscotch} genotyping data}

    \begin{tabular}{cccc}\\\toprule
    \textbf{Comparison} & \textbf{Region 1} & \textbf{Region 2} & \textbf{\emph{Hopscotch}} \\\midrule
    EjuA \& EjuB & 0     & 0     & 0 \\
    EjuA \& MSA & 0.326 & 0.328 & 0.186 \\
    EjuA \& SLO & 0.416 & 0.258 & 0.280 \\
    EjuB \& MSA & 0.397 & 0.365 & 0.188 \\
    EjuB \& SLO & 0.512  & 0.290 & 0.280 \\
    MSA \& SLO & 0.007  & 0 & 0.016 \\\bottomrule
    \end{tabular}
  \label{Table1Fst}
\end{table}

% Table S1. Parv and mex samples 
\begin{table}[htbp]
\rowcolors{2}{gray!25}{white}
  \centering
  \caption{Accessions of \emph{Zea mays} ssp. \emph{mexicana} (RIMME) and \emph{Zea mays} ssp. \emph{parviglumis} (RIMPA) sampled. RIHY is a \emph{Z. mays} ssp. \emph{parviglumis} and \emph{Zea mays} ssp. \emph{mays} hybrid. }
    \begin{tabular}{ccccc}\\\toprule
    \textbf{Accession} & \textbf{USDA Accession ID} & \textbf{Locality} & \textbf{Number Alleles Sampled} & \textbf{\emph{Hopscotch} Frequency} & \textbf{No \emph{Hopscotch} Frequency} \\
    RIHY0009 & N/A   & N/A   & 2     & 0.5   & 0.5 \\
    RIMME0006 & 566673 & Durango & 2     & 0     & 1 \\
    RIMME0007 & 566680 & Guanajuato & 2     & 0     & 1 \\
    RIMME0008 & 566681 & Michoacan & 2     & 0     & 1 \\
    RIMME0009 & 566682 & Distrito Federal & 2     & 0     & 1 \\
    RIMME0011 & 566685 & Mexico & 2     & 0     & 1 \\
    RIMME0014 & 714151 & Breeders line; Puga: 11066 & 6     & 0     & 1 \\
    RIMME0017 & 699874 & Ayotlan & 8     & 0     & 1 \\
    RIMME0021 & N/A   & El Porvenir & 69    & 0.173913043 & 0.826086957 \\
    RIMME0026 & N/A   & Opopeo & 42    & 0.071428571 & 0.928571429 \\
    RIMME0028 & N/A   & Puruandiro & 28    & 0.035714286 & 0.964285714 \\
    RIMME0029 & N/A   & Ixtlan & 35    & 0     & 1 \\
    RIMME0030 & N/A   & San Pedro & 27    & 0     & 1 \\
    RIMME0031 & N/A   & Tenango del Aire & 25    & 0.08  & 0.92 \\
    RIMME0032 & N/A   & Nabogame & 24    & 0     & 1 \\
    RIMME0033 & N/A   & Puerta Encantada & 25    & 0     & 1 \\
    RIMME0034 & N/A   & Santa Clara & 23    & 0     & 1 \\
    RIMME0035 & N/A   & Xochimilco & 25    & 0     & 1 \\
    RIMPA0001 & 87168 & El Salado & 4     & 0     & 1 \\
    RIMPA0003 & 87171 & Mazatlan & 8     & 0.125 & 0.875 \\
    RIMPA0017 & 87200 & N/A   & 4     & 0     & 1 \\
    RIMPA0019 & 87213 & El Salado & 2     & 0.5   & 0.5 \\
    RIMPA0029 & 87244 & N/A   & 2     & 0.5   & 0.5 \\
    RIMPA0031 & 87249 & N/A   & 2     & 0.5   & 0.5 \\
    RIMPA0035 & 87288 & Jalisco & 4     & 0     & 1 \\
    RIMPA0040 & 288185 & Mexico & 4     & 0     & 1 \\
    RIMPA0042 & 288187 & Guerrero & 4     & 0.25  & 0.75 \\
    RIMPA0043 & 288188 & Guerrero & 4     & 0     & 1 \\
    RIMPA0045 & 288193 & Guerrero & 4     & 0     & 1 \\
    RIMPA0055 & 714152 & Breeders line & 2     & 0     & 1 \\
    RIMPA0056 & 714153 & Breeders line & 2     & 0.5   & 0.5 \\
    RIMPA0057 & 714154 & Breeders line & 2     & 0.5   & 0.5 \\
    RIMPA0058 & N/A   & N/A   & 4     & 0.5   & 0.5 \\
    RIMPA0059 & N/A   & N/A   & 4     & 1     & 0 \\
    RIMPA0060 & 714157 & Breeders line: CIMMYT 11355 & 2     & 0     & 1 \\
    RIMPA0061 & 714158 & Breeders line: USDA PI566686 & 4     & 0.5   & 0.5 \\
    RIMPA0062 & 714159 & Breeders line & 4     & 0.5   & 0.5 \\
    RIMPA0063 & 714160 & Breeders line & 4     & 0     & 1 \\
    RIMPA0064 & 714161 & Breeders line & 3     & 0     & 1 \\
    RIMPA0065 & 714162 & Breeders line & 4     & 0.25  & 0.75 \\
    RIMPA0068 & 699861 & Jalisco, Mexico & 16    & 0     & 1 \\
    RIMPA0069 & 699862 & Ixtlan & 14    & 0.142857143 & 0.857142857 \\
    RIMPA0070 & 699863 & Benito Jaurez & 16    & 0     & 1 \\
    RIMPA0071 & 699864 & Tuzantla & 28    & 0     & 1 \\
    RIMPA0072 & 699865 & Tiquicheo & 16    & 0     & 1 \\
    RIMPA0073 & 699866 & Tiquicheo & 16    & 0.125 & 0.875 \\
    RIMPA0074 & 699867 & Huetamo & 12    & 0     & 1 \\
    RIMPA0075 & 699868 & Huetamo & 2     & 0     & 1 \\
    RIMPA0076 & 699869 & Huetamo & 4     & 0     & 1 \\
    RIMPA0077 & 699870 & Caracuaro & 2     & 0     & 1 \\
    RIMPA0078 & 699871 & Caracuaro & 2     & 0.5   & 0.5 \\
    RIMPA0079 & 699872 & Villa Madero & 14    & 0     & 1 \\
    RIMPA0080 & 699873 & Guachinango & 12    & 0     & 1 \\
    RIMPA0081 & 699875 & Ameca & 16    & 0     & 1 \\
    RIMPA0083 & 699877 & Tepoztlan & 14    & 0     & 1 \\
    RIMPA0084 & 699878 & Tepoztlan & 16    & 0     & 1 \\
    RIMPA0085 & 699879 & Miahuatlan & 16    & 0     & 1 \\
    RIMPA0086 & 699880 & Miahuatlan & 16    & 0.0625 & 0.9375 \\
    RIMPA0087 & 699881 & Tecoanapa & 24    & 0     & 1 \\
    RIMPA0089 & 699883 & Guerrero & 12    & 0     & 1 \\
    RIMPA0090 & 699884 & Guerrero & 10    & 0     & 1 \\
    RIMPA0091 & 699885 & Guerrero & 16    & 0     & 1 \\
    RIMPA0092 & 699886 & Guerrero & 10    & 0     & 1 \\
    RIMPA0093 & 699887 & Guerrero & 26    & 0.076923077 & 0.923076923 \\
    RIMPA0094 & 699888 & Guerrero & 2     & 0     & 1 \\
    RIMPA0095 & 699889 & Guerrero & 4     & 0     & 1 \\
    RIMPA0096 & 699890 & Guerrero & 26    & 0.038461538 & 0.961538462 \\
    RIMPA0097 & 699891 & Guerrero & 6     & 0     & 1 \\
    RIMPA0098 & 699892 & Guerrero & 4     & 0     & 1 \\
    RIMPA0099 & 699893 & Guerrero & 4     & 0     & 1 \\
    RIMPA0100 & 699894 & Guerrero & 6     & 0     & 1 \\
    RIMPA0101 & 699895 & Guerrero & 2     & 0     & 1 \\
    RIMPA0103 & 699897 & Guerrero & 2     & 0     & 1 \\
    RIMPA0104 & 699898 & Guerrero & 22    & 0.090909091 & 0.909090909 \\
    RIMPA0105 & 699899 & Guerrero & 6     & 0     & 1 \\
    RIMPA0106 & 699900 & Guerrero & 6     & 0.333333333 & 0.666666667 \\
    RIMPA0107 & 699901 & Guerrero & 4     & 0     & 1 \\
    RIMPA0108 & 699902 & Guerrero & 6     & 0     & 1 \\
    RIMPA0109 & 699903 & Michoacan & 4     & 0.25  & 0.75 \\
    RIMPA0110 & 699904 & Michoacan & 2     & 0     & 1 \\
    RIMPA0111 & 699905 & Michoacan & 4     & 0     & 1 \\
    RIMPA0112 & 699906 & Michoacan & 4     & 0.25  & 0.75 \\
    RIMPA0114 & 699908 & Michoacan & 6     & 0.166666667 & 0.833333333 \\
    RIMPA0116 & 699910 & Mexico & 2     & 0     & 1 \\
    RIMPA0117 & 699911 & Mexico & 4     & 0     & 1 \\
    RIMPA0118 & 699912 & Mexico & 6     & 0.166666667 & 0.833333333 \\
    RIMPA0119 & 699913 & Mexico & 2     & 0     & 1 \\
    RIMPA0120 & 699914 & Mexico & 1     & 1     & 0 \\
    RIMPA0121 & 699915 & Mexico & 2     & 0     & 1 \\
    RIMPA0128 & 699922 & Mexico & 2     & 0.5   & 0.5 \\
    RIMPA0129 & 699923 & Michoacan & 2     & 0.5   & 0.5 \\
    RIMPA0135 & 699929 & Nayarit & 24    & 0     & 1 \\
    RIMPA0138 & 699932 & Jalisco & 2     & 0.5   & 0.5 \\
    RIMPA0139 & 699933 & Jalisco & 1     & 1     & 0 \\
    RIMPA0142 & 699936 & Colima & 18    & 0.444444444 & 0.555555556 \\
    RIMPA0144 & 699938 & Jalisco & 2     & 1     & 0 \\
    RIMPA0145 & 699939 & Michoacan & 1     & 1     & 0 \\
    RIMPA0147 & 699941 & Jalisco & 1     & 1     & 0 \\
    RIMPA0155 & N/A   & Jalisco & 73    & 0.01369863 & 0.98630137 \\
    RIMPA0156 & N/A   & Jalisco & 20    & 0     & 1 \\
    RIMPA0157 & N/A   & Jalisco & 58    & 0.344827586 & 0.655172414 \\
    RIMPA0158 & N/A   & Jalisco & 64    & 0.53125 & 0.46875 \\
    RIMPA0159 & N/A   & Jalisco & 26    & 0     & 1 \\
    RIMPA0162 & 21785 & N/A   & 4     & 0     & 1 \\\bottomrule
    \end{tabular}
  \label{TableS1}
\end{table}

% Table S2. Mays sampling
\begin{table}[htbp]
\rowcolors{2}{gray!25}{white}
  \centering

  \centering
 \caption{\emph{Zea mays} ssp. \emph{mays} (RIMMA) sampled for genotyping}
    \begin{tabular}{ccc}
 %   \toprule
    Accession & Number of alleles sampled & \emph{Hopscotch} Frequency \\
 %   \midrule
    RIMMA0066 & 2     & 1 \\
    RIMMA0075 & 2     & 1 \\
    RIMMA0077 & 2     & 1 \\
    RIMMA0079 & 2     & 1 \\
    RIMMA0081 & 2     & 1 \\
    RIMMA0084 & 2     & 1 \\
    RIMMA0086 & 2     & 1 \\
    RIMMA0088 & 2     & 1 \\
    RIMMA0089 & 2     & 1 \\
    RIMMA0090 & 2     & 1 \\
    RIMMA0092 & 4     & 1 \\
    RIMMA0094 & 4     & 1 \\
    RIMMA0097 & 2     & 1 \\
    RIMMA0099 & 2     & 1 \\
    RIMMA0100 & 2     & 1 \\
    RIMMA0101 & 2     & 1 \\
    RIMMA0104 & 2     & 1 \\
    RIMMA0108 & 2     & 1 \\
    RIMMA0111 & 6     & 1 \\
    RIMMA0115 & 2     & 1 \\
    RIMMA0117 & 2     & 1 \\
    RIMMA0130 & 2     & 1 \\
    RIMMA0133 & 2     & 1 \\
    RIMMA0134 & 2     & 1 \\
    RIMMA0135 & 2     & 1 \\
    RIMMA0142 & 2     & 0.5 \\
    RIMMA0143 & 4     & 1 \\
    RIMMA0146 & 4     & 1 \\
    RIMMA0149 & 2     & 1 \\
    RIMMA0152 & 2     & 1 \\
    RIMMA0153 & 2     & 1 \\
    RIMMA0154 & 2     & 1 \\
    RIMMA0155 & 2     & 1 \\
    RIMMA0156 & 2     & 1 \\
    RIMMA0157 & 2     & 1 \\
    RIMMA0158 & 2     & 1 \\
    RIMMA0159 & 2     & 1 \\
    RIMMA0160 & 2     & 1 \\
    RIMMA0162 & 2     & 1 \\
    RIMMA0166 & 2     & 1 \\
    RIMMA0167 & 2     & 1 \\
    RIMMA0168 & 2     & 1 \\
    RIMMA0169 & 2     & 1 \\
    RIMMA0172 & 2     & 1 \\
    RIMMA0174 & 4     & 1 \\
    RIMMA0177 & 2     & 1 \\
    RIMMA0178 & 2     & 1 \\
    RIMMA0179 & 2     & 1 \\
    RIMMA0181 & 2     & 1 \\
    RIMMA0183 & 2     & 1 \\
    RIMMA0184 & 2     & 1 \\
    RIMMA0186 & 2     & 1 \\
    RIMMA0187 & 1     & 1 \\
    RIMMA0188 & 2     & 1 \\
    RIMMA0195 & 2     & 1 \\
    RIMMA0196 & 2     & 1 \\
    RIMMA0197 & 2     & 1 \\
    RIMMA0198 & 2     & 1 \\
    RIMMA0199 & 2     & 1 \\
    RIMMA0200 & 2     & 1 \\
    RIMMA0202 & 2     & 1 \\
    RIMMA0203 & 2     & 1 \\
    RIMMA0206 & 2     & 1 \\
    RIMMA0208 & 2     & 1 \\
    RIMMA0209 & 2     & 1 \\
    RIMMA0210 & 2     & 1 \\
    RIMMA0212 & 2     & 1 \\
    RIMMA0213 & 2     & 1 \\
    RIMMA0214 & 2     & 1 \\
    RIMMA0217 & 2     & 1 \\
    RIMMA0218 & 2     & 1 \\
    RIMMA0220 & 2     & 1 \\
    RIMMA0221 & 2     & 1 \\
    RIMMA0222 & 2     & 1 \\
    RIMMA0223 & 2     & 1 \\
    RIMMA0226 & 2     & 1 \\
    RIMMA0227 & 2     & 1 \\
    RIMMA0228 & 2     & 1 \\
    RIMMA0229 & 2     & 1 \\
    RIMMA0230 & 2     & 1 \\
    RIMMA0232 & 2     & 1 \\
    RIMMA0233 & 2     & 1 \\
    RIMMA0235 & 2     & 0.5 \\
    RIMMA0242 & 2     & 1 \\
    RIMMA0243 & 2     & 1 \\
    RIMMA0247 & 4     & 1 \\
    RIMMA0248 & 2     & 1 \\
    RIMMA0249 & 2     & 1 \\
    RIMMA0252 & 2     & 1 \\
    RIMMA0253 & 2     & 1 \\
    RIMMA0254 & 2     & 1 \\
    RIMMA0256 & 2     & 1 \\
    RIMMA0257 & 2     & 1 \\
    RIMMA0258 & 2     & 1 \\
    RIMMA0259 & 2     & 1 \\
    RIMMA0260 & 2     & 1 \\
    RIMMA0262 & 2     & 1 \\
    RIMMA0263 & 2     & 1 \\
    RIMMA0264 & 2     & 1 \\
    RIMMA0265 & 2     & 1 \\
    RIMMA0268 & 2     & 1 \\
    RIMMA0269 & 2     & 1 \\
    RIMMA0270 & 2     & 1 \\
    RIMMA0272 & 2     & 1 \\
    RIMMA0275 & 2     & 1 \\
    RIMMA0276 & 2     & 1 \\
    RIMMA0277 & 2     & 1 \\
    RIMMA0279 & 2     & 1 \\
    RIMMA0280 & 2     & 1 \\
    RIMMA0283 & 2     & 1 \\
    RIMMA0285 & 2     & 1 \\
    RIMMA0288 & 2     & 1 \\
    RIMMA0290 & 2     & 1 \\
    RIMMA0291 & 2     & 1 \\
    RIMMA0292 & 2     & 1 \\
    RIMMA0293 & 2     & 1 \\
    RIMMA0298 & 2     & 1 \\
    RIMMA0302 & 2     & 1 \\
    RIMMA0305 & 2     & 1 \\
    RIMMA0310 & 2     & 1 \\
    RIMMA0312 & 2     & 1 \\
    RIMMA0320 & 2     & 1 \\
    RIMMA0322 & 2     & 1 \\
    RIMMA0324 & 2     & 1 \\
    RIMMA0334 & 2     & 1 \\
    RIMMA0336 & 2     & 1 \\
    RIMMA0337 & 2     & 1 \\
    RIMMA0338 & 2     & 1 \\
    RIMMA0339 & 2     & 1 \\
    RIMMA0340 & 2     & 1 \\
    RIMMA0341 & 2     & 1 \\
    RIMMA0342 & 2     & 1 \\
    RIMMA0344 & 2     & 1 \\
    RIMMA0346 & 2     & 1 \\
    RIMMA0347 & 2     & 1 \\
    RIMMA0350 & 2     & 1 \\
    RIMMA0351 & 2     & 1 \\
    RIMMA0352 & 2     & 1 \\
    RIMMA0357 & 2     & 1 \\
    RIMMA0360 & 18    & 1 \\
    RIMMA0362 & 8     & 1 \\
    RIMMA0363 & 16    & 1 \\
    RIMMA0370 & 17    & 1 \\
    RIMMA0372 & 24    & 1 \\
    RIMMA0373 & 22    & 1 \\
    RIMMA0408 & 2     & 1 \\
    RIMMA0411 & 1     & 1 \\
    RIMMA0413 & 1     & 1 \\
    RIMMA0414 & 1     & 1 \\
    RIMMA0424 & 8     & 0.9 \\
    RIMMA0432 & 1     & 1 \\
    RIMMA0434 & 2     & 0.5 \\
    RIMMA0435 & 1     & 1 \\
    RIMMA0442 & 2     & 1 \\
    RIMMA0443 & 2     & 1 \\
    RIMMA0444 & 2     & 1 \\
    RIMMA0445 & 2     & 1 \\
    RIMMA0446 & 2     & 1 \\
    RIMMA0447 & 2     & 1 \\
    RIMMA0448 & 2     & 1 \\
    RIMMA0449 & 2     & 1 \\
    RIMMA0450 & 2     & 1 \\
    RIMMA0451 & 2     & 1 \\
    RIMMA0452 & 2     & 1 \\
    RIMMA0453 & 2     & 1 \\
    RIMMA0454 & 2     & 1 \\
    RIMMA0455 & 2     & 1 \\
    RIMMA0456 & 2     & 1 \\
    RIMMA0457 & 2     & 1 \\
    RIMMA0458 & 2     & 1 \\
    RIMMA0459 & 2     & 1 \\
    RIMMA0483 & 2     & 1 \\
    RIMMA0490 & 2     & 1 \\
    RIMMA0515 & 2     & 1 \\
    RIMMA0537 & 2     & 1 \\
    RIMMA0550 & 2     & 1 \\
    RIMMA0553 & 2     & 1 \\
    RIMMA0559 & 2     & 1 \\
    RIMMA0561 & 2     & 1 \\
    RIMMA0562 & 2     & 1 \\
    RIMMA0571 & 2     & 1 \\
    RIMMA0572 & 2     & 1 \\
    RIMMA0577 & 2     & 1 \\
    RIMMA0579 & 2     & 1 \\
    RIMMA0590 & 2     & 1 \\
    RIMMA0591 & 2     & 1 \\
    RIMMA0592 & 2     & 1 \\
    RIMMA0593 & 2     & 1 \\
    RIMMA0594 & 2     & 1 \\
    RIMMA0595 & 2     & 1 \\
    RIMMA0596 & 2     & 1 \\
    RIMMA0597 & 2     & 1 \\
    RIMMA0598 & 2     & 1 \\
    RIMMA0599 & 2     & 1 \\
    RIMMA0600 & 2     & 1 \\
    RIMMA0601 & 2     & 1 \\
    RIMMA0602 & 2     & 1 \\
    RIMMA0603 & 2     & 1 \\
    RIMMA0604 & 2     & 1 \\
    RIMMA0605 & 2     & 1 \\
    RIMMA0606 & 2     & 1 \\
    RIMMA0607 & 2     & 1 \\
    RIMMA0608 & 2     & 1 \\
    RIMMA0609 & 2     & 1 \\
    RIMMA0610 & 2     & 1 \\
    RIMMA0611 & 2     & 1 \\
    RIMMA0612 & 2     & 1 \\
    RIMMA0613 & 2     & 1 \\
    RIMMA0622 & 2     & 0.5 \\
    RIMMA0624 & 1     & 1 \\
    RIMMA0629 & 2     & 0.5 \\
    RIMMA0631 & 2     & 0.5 \\
    RIMMA0659 & 2     & 1 \\
    RIMMA0660 & 2     & 1 \\
    RIMMA0678 & 1     & 1 \\
    RIMMA0681 & 2     & 1 \\
    RIMMA0683 & 2     & 1 \\
    RIMMA0684 & 2     & 1 \\
    RIMMA0685 & 4     & 1 \\
    RIMMA0693 & 2     & 1 \\
    RIMMA0694 & 2     & 1 \\
    RIMMA0695 & 1     & 1 \\
    RIMMA0699 & 2     & 1 \\
    RIMMA0704 & 2     & 1 \\
    RIMMA0706 & 2     & 1 \\
    RIMMA0711 & 1     & 1 \\
    RIMMA0713 & 2     & 1 \\
    RIMMA0715 & 2     & 1 \\
    RIMMA0717 & 2     & 1 \\
    RIMMA0718 & 2     & 1 \\
    RIMMA0723 & 2     & 1 \\
    RIMMA0724 & 2     & 1 \\
    RIMMA0725 & 1     & 1 \\
    RIMMA0728 & 1     & 1 \\
    RIMMA0732 & 2     & 1 \\
    RIMMA0734 & 2     & 1 \\
    RIMMA0738 & 2     & 1 \\
    RIMMA0739 & 2     & 1 \\
    RIMMA0744 & 2     & 1 \\
    RIMMA0747 & 1     & 1 \\
    RIMMA0748 & 2     & 1 \\
    RIMMA0749 & 2     & 1 \\
    RIMMA0750 & 2     & 0.5 \\
    RIMMA0751 & 1     & 1 \\
    RIMMA0752 & 1     & 1 \\
    RIMMA0753 & 2     & 1 \\
    RIMMA0755 & 2     & 1 \\
    \bottomrule
    \end{tabular}
  \label{TableS2Mays}
\end{table}

% Table S3. PCs for BayEnv
\begin{table}[htbp]
  \centering
  \caption{Variables and rotations used for the 6 principal components used for BayEnv calculations and their corresponding Bayes Factors. Modified from \citet{Pyhajarvi2013}.}
    \begin{tabular}{rrrrrrrrrrrr}
    \toprule
    \textbf{PC1} & \textbf{} & \textbf{PC2} & \textbf{} & \textbf{PC3} & \textbf{} & \textbf{PC4} & \textbf{} & \textbf{PC5} & \textbf{} & \textbf{PC6} & \textbf{} \\
    \midrule
    \textbf{Var} & \textbf{Rot} & \textbf{Var} & \textbf{Rot} & \textbf{Var} & \textbf{Rot} & \textbf{Var} & \textbf{Rot} & \textbf{Var} & \textbf{Rot} & \textbf{Var} & \textbf{Rot} \\
    bio1  & 0.146 & bio4  & 0.244 & prec7 & 0.287 & ts\_clay & 0.41  & bio2  & 0.38  & bio2  & 0.412 \\
    tmean11 & 0.146 & bio3  & 0.241 & prec8 & 0.276 & v\_mod & 0.359 & sq4   & 0.328 & x\_mod & 0.365 \\
    tmean12 & 0.145 & bio7  & 0.241 & prec11 & 0.262 & ts\_sand & 0.329 & ts\_loam & 0.289 & sq7   & 0.332 \\
    bio11 & 0.145 & prec6 & 0.237 & bio13 & 0.247 & bio15 & 0.272 & ts\_sand & 0.266 & bio7  & 0.312 \\
    tmax12 & 0.145 & sq7   & 0.218 & prec1 & 0.246 & prec4 & 0.259 & sq7   & 0.231 & v\_mod & 0.307 \\
    tmin5 & 0.145 & prec9 & 0.217 & bio16 & 0.242 & x\_mod & 0.244 & bio18 & 0.213 & prec11 & 0.23 \\
    tmean1 & 0.145 & sq3   & 0.207 & prec12 & 0.24  & prec3 & 0.226 & bio13 & 0.207 & bio18 & 0.22 \\
    tmean2 & 0.145 & prec12 & 0.207 & bio19 & 0.238 & sq3   & 0.21  & prec11 & 0.183 & sq3   & 0.176 \\
    tmin4 & 0.145 & bio12 & 0.204 & bio12 & 0.231 & prec5 & 0.21  & bio7  & 0.17  & sq4   & 0.153 \\
    tmax1 & 0.145 & bio19 & 0.196 & prec2 & 0.222 & prec7 & 0.19  & bio16 & 0.163 & ts\_sand & 0.153 \\
    tmean4 & 0.145 & prec2 & 0.188 & bio18 & 0.221 & sq4   & 0.186 & bio4  & 0.157 & bio4  & 0.148 \\
    tmin11 & 0.144 & prec1 & 0.185 & sq4   & 0.2   & bio3  & 0.185 & bio12 & 0.156 & prec7 & 0.139 \\
    tmax11 & 0.144 & prec10 & 0.184 & prec9 & 0.18  & bio18 & 0.178 & bio3  & 0.155 & tmax3 & 0.127 \\
    tmin12 & 0.144 & bio16 & 0.183 & prec10 & 0.171 & sq7   & 0.132 & prec6 & 0.154 & tmax4 & 0.126 \\
    tmin2 & 0.144 & prec8 & 0.17  & prec5 & 0.161 & bio14 & 0.116 & x\_mod & 0.152 & bio5  & 0.118 \\
    tmean5 & 0.144 & prec5 & 0.165 & prec4 & 0.154 & bio13 & 0.099 & prec9 & 0.144 & tmax5 & 0.115 \\
    tmean10 & 0.144 & bio14 & 0.158 & sq3   & 0.147 & bio16 & 0.095 & prec8 & 0.143 & bio3  & 0.107 \\
    bio6  & 0.144 & bio13 & 0.151 & bio2  & 0.143 & prec8 & 0.09  & v\_mod & 0.142 & ts\_loam & 0.095 \\
    tmax2 & 0.144 & bio17 & 0.149 & bio17 & 0.129 & bio7  & 0.077 & bio15 & 0.136 & ts\_clay & 0.078 \\
    tmean3 & 0.144 & prec3 & 0.144 & ts\_loam & 0.127 & bio4  & 0.075 & prec7 & 0.112 & tmin9 & 0.072 \\
    tmin1 & 0.143 & ts\_clay & 0.141 & v\_mod & 0.123 & bio2  & 0.074 & prec4 & 0.108 & tmin8 & 0.071 \\
    tmin10 & 0.143 & bio2  & 0.129 & prec3 & 0.113 & prec2 & 0.074 & bio14 & 0.096 & prec9 & 0.07 \\
    Altitude & 0.143 & prec7 & 0.108 & x\_mod & 0.111 & bio19 & 0.068 & tmax7 & 0.093 & tmin10 & 0.069 \\
    bio9  & 0.143 & tmax6 & 0.107 & bio14 & 0.099 & prec12 & 0.056 & tmax8 & 0.092 & tmin12 & 0.069 \\
    tmin3 & 0.143 & x\_mod & 0.106 & bio4  & 0.07  & ts\_loam & 0.053 & prec1 & 0.091 & tmin7 & 0.066 \\
    bio10 & 0.142 & bio15 & 0.098 & tmax3 & 0.067 & tmax12 & 0.047 & prec2 & 0.086 & tmean4 & 0.066 \\
    tmax10 & 0.142 & ts\_loam & 0.088 & ts\_clay & 0.065 & bio17 & 0.047 & tmin11 & 0.086 & tmax2 & 0.066 \\
    tmax3 & 0.142 & tmean6 & 0.085 & bio15 & 0.056 & bio9  & 0.043 & prec5 & 0.082 & tmax6 & 0.064 \\
    tmax4 & 0.142 & tmin7 & 0.082 & tmax2 & 0.055 & tmax8 & 0.042 & bio17 & 0.082 & tmean3 & 0.064 \\
    tmin6 & 0.142 & bio5  & 0.082 & tmean3 & 0.052 & tmax1 & 0.041 & tmin12 & 0.08  & bio9  & 0.058 \\
    tmean9 & 0.141 & tmean7 & 0.081 & ts\_sand & 0.05  & tmax5 & 0.039 & prec3 & 0.078 & tmin11 & 0.058 \\
    tmin9 & 0.141 & prec4 & 0.08  & prec6 & 0.048 & tmax7 & 0.039 & tmax9 & 0.078 & prec3 & 0.056 \\
    tmean8 & 0.141 & tmax7 & 0.079 & sq7   & 0.048 & prec10 & 0.038 & tmin1 & 0.077 & tmean5 & 0.056 \\
    bio8  & 0.14  & bio8  & 0.079 & tmin7 & 0.046 & Altitude & 0.037 & tmin10 & 0.074 & bio10 & 0.051 \\
    tmean6 & 0.14  & tmax9 & 0.077 & bio3  & 0.044 & tmax10 & 0.035 & bio6  & 0.071 & bio15 & 0.047 \\
    tmean7 & 0.14  & tmean8 & 0.076 & tmax4 & 0.043 & tmax2 & 0.033 & prec12 & 0.067 & tmin1 & 0.045 \\
    tmin8 & 0.14  & tmin8 & 0.076 & bio7  & 0.042 & tmax9 & 0.03  & tmin2 & 0.061 & prec12 & 0.045 \\
    tmax5 & 0.14  & tmax5 & 0.074 & tmax1 & 0.036 & tmean12 & 0.029 & tmin6 & 0.061 & bio17 & 0.045 \\
    tmax9 & 0.139 & tmax8 & 0.074 & tmin3 & 0.035 & tmax11 & 0.027 & tmax6 & 0.059 & tmean10 & 0.043 \\
    tmax8 & 0.139 & tmean9 & 0.072 & bio9  & 0.035 & tmean1 & 0.027 & tmin3 & 0.052 & prec4 & 0.043 \\
    bio5  & 0.139 & bio18 & 0.07  & tmin8 & 0.034 & bio5  & 0.026 & bio8  & 0.046 & Altitude & 0.041 \\
    tmax7 & 0.139 & v\_mod & 0.066 & tmean4 & 0.031 & tmean5 & 0.026 & tmean11 & 0.041 & bio6  & 0.039 \\
    tmin7 & 0.138 & tmin9 & 0.066 & tmean2 & 0.031 & bio11 & 0.025 & tmean1 & 0.04  & tmin6 & 0.038 \\
    tmax6 & 0.135 & bio10 & 0.065 & tmax12 & 0.03  & tmean2 & 0.022 & tmin9 & 0.04  & tmean12 & 0.035 \\
    bio17 & 0.107 & tmax10 & 0.063 & tmean7 & 0.028 & tmin6 & 0.022 & tmean12 & 0.04  & tmean9 & 0.035 \\
    bio14 & 0.095 & tmin6 & 0.061 & tmin9 & 0.023 & prec1 & 0.022 & tmax10 & 0.038 & prec5 & 0.033 \\
    prec3 & 0.094 & prec11 & 0.056 & Altitude & 0.021 & tmean7 & 0.02  & tmin5 & 0.038 & tmean8 & 0.03 \\
    bio15 & 0.092 & ts\_sand & 0.054 & tmax11 & 0.019 & tmean8 & 0.02  & bio11 & 0.037 & tmin2 & 0.03 \\
    prec4 & 0.088 & sq4   & 0.049 & tmin4 & 0.018 & tmean9 & 0.019 & tmean8 & 0.035 & tmax1 & 0.03 \\
    prec10 & 0.084 & tmin3 & 0.048 & tmax5 & 0.017 & prec11 & 0.019 & tmean7 & 0.033 & prec1 & 0.025 \\
    ts\_sand & 0.073 & tmean10 & 0.047 & tmin10 & 0.017 & tmean11 & 0.018 & tmean2 & 0.03  & bio13 & 0.024 \\
    ts\_loam & 0.072 & tmean5 & 0.044 & tmin6 & 0.017 & tmax4 & 0.016 & bio19 & 0.029 & tmean11 & 0.023 \\
    prec6 & 0.069 & tmax4 & 0.042 & tmean6 & 0.017 & tmean10 & 0.016 & bio9  & 0.027 & tmax7 & 0.023 \\
    prec9 & 0.063 & tmax11 & 0.041 & tmax6 & 0.017 & tmin5 & 0.015 & tmin7 & 0.026 & tmean7 & 0.023 \\
    prec2 & 0.063 & bio6  & 0.038 & tmin12 & 0.017 & tmean6 & 0.015 & tmin4 & 0.025 & prec10 & 0.022 \\
    bio4  & 0.06  & tmin1 & 0.038 & bio11 & 0.016 & prec6 & 0.014 & tmean5 & 0.023 & prec2 & 0.02 \\
    prec5 & 0.058 & tmin2 & 0.037 & tmin11 & 0.015 & bio6  & 0.012 & tmean3 & 0.022 & bio8  & 0.018 \\
    bio13 & 0.055 & Altitude & 0.036 & tmin1 & 0.013 & tmin1 & 0.012 & tmin8 & 0.022 & tmean2 & 0.017 \\
    x\_mod & 0.053 & tmin10 & 0.034 & tmean5 & 0.013 & tmin12 & 0.011 & tmean10 & 0.021 & bio16 & 0.016 \\
    prec1 & 0.05  & bio9  & 0.032 & tmean8 & 0.013 & tmin2 & 0.011 & tmean9 & 0.019 & tmean6 & 0.014 \\
    bio16 & 0.049 & tmean1 & 0.029 & bio6  & 0.012 & bio1  & 0.011 & tmean4 & 0.019 & bio19 & 0.012 \\
    bio7  & 0.043 & bio1  & 0.028 & tmean9 & 0.01  & tmin4 & 0.009 & ts\_clay & 0.018 & tmax11 & 0.012 \\
    bio19 & 0.042 & tmean3 & 0.026 & tmean1 & 0.01  & tmax6 & 0.008 & prec10 & 0.015 & tmax10 & 0.011 \\
    bio12 & 0.042 & tmean2 & 0.024 & bio10 & 0.01  & tmin11 & 0.007 & bio1  & 0.014 & tmax8 & 0.01 \\
    bio3  & 0.04  & bio11 & 0.022 & tmax8 & 0.009 & bio8  & 0.007 & tmax4 & 0.014 & tmean1 & 0.009 \\
    prec7 & 0.034 & tmean11 & 0.019 & tmax7 & 0.009 & bio10 & 0.007 & bio10 & 0.011 & bio14 & 0.009 \\
    prec12 & 0.033 & tmin12 & 0.019 & bio5  & 0.008 & tmin9 & 0.006 & tmax5 & 0.009 & prec8 & 0.007 \\
    bio2  & 0.033 & tmax1 & 0.017 & tmin5 & 0.008 & prec9 & 0.005 & bio5  & 0.008 & bio11 & 0.007 \\
    prec8 & 0.032 & tmean4 & 0.016 & tmin2 & 0.007 & bio12 & 0.004 & tmax11 & 0.008 & tmin5 & 0.006 \\
    sq4   & 0.026 & tmin5 & 0.013 & bio1  & 0.007 & tmax3 & 0.003 & tmax12 & 0.008 & tmax9 & 0.005 \\
    sq3   & 0.024 & tmean12 & 0.012 & tmax10 & 0.006 & tmean4 & 0.003 & tmax3 & 0.007 & bio12 & 0.004 \\
    v\_mod & 0.015 & tmax2 & 0.011 & tmean10 & 0.006 & tmin3 & 0.002 & sq3   & 0.007 & tmin3 & 0.004 \\
    ts\_clay & 0.012 & tmin4 & 0.01  & tmean12 & 0.005 & tmin7 & 0.002 & tmax2 & 0.006 & prec6 & 0.004 \\
    sq7   & 0.004 & tmax3 & 0.003 & bio8  & 0.004 & tmean3 & 0.002 & Altitude & 0.004 & tmin4 & 0.003 \\
    prec11 & 0.004 & tmax12 & 0.002 & tmax9 & 0.003 & tmin10 & 0.002 & tmean6 & 0.002 & bio1  & 0.002 \\
    bio18 & 0.001 & tmin11 & 0.001 & tmean11 & 0.001 & tmin8 & 0     & tmax1 & 0.001 & tmax12 & 0.002 \\
    \textbf{Bayes Factors} &       &       &       &       &       &       &       &       &       &       &  \\
    \multicolumn{2}{c}{0.77165} & \multicolumn{2}{c}{0.51652} & \multicolumn{2}{c}{0.21120} & \multicolumn{2}{c}{0.24063} & \multicolumn{2}{c}{1.96010} & \multicolumn{2}{c}{0.24594} \\
    \bottomrule
    \end{tabular}%
  \label{TableS3BayEnv}%
\end{table}%

% Table S4 Diversity stats in tb1 region based on SNP 50 genotyping data
\begin{table}[htbp]
  \centering
  \caption{Diversity in the \emph{tb1} region based on the maize SNP50 genotyping data}
    \begin{tabular}{ccccc}
    \toprule
    \textbf{Population} & \textbf{\# seg sites} & \textbf{$\theta_{\subscript{\pi}}$ / bp} & \textbf{$\theta_{\subscript{W}}$ / bp} & \textbf{Tajima?s D} \\
    \midrule
    Ejutla A & 4     & 0.15217 & 0.11902 & 0.76191 \\
    Ejutla B & 5     & 0.15258 & 0.14877 & 0.07412 \\
    La Mesa & 3     & 0.12802 & 0.08926 & 1.09209 \\
    San Lorenzo & 3     & 0.09098 & 0.08926 & 0.04845 \\
    \bottomrule
    \end{tabular}%
  \label{TableS4DiversitySNP}%
\end{table}%

\end{document}