### 2014-05-29 Supplement_mbh.tex
final formatting for supplement

%Online Supplemental Materials
%Authors may wish to augment their manuscripts with online supplemental materials (e.g., large data sets, three-dimensional reconstructions, simulations, real-time movies, color photographs). Upload these appendices as separate files with the initial manuscript submission. Include a header on each file using this format: Smith et al.?American Journal of Botany 99(#): ###-###. 2012. ? Data Supplement S1 ? Page 1?. Name online supplements Appendix S1, Appendix S2, etc, in the order in which they appear in the text, regardless of whether they are tables, figures, text, other media, or a combination thereof. In the manuscript, after the mention of an online appendix, include the following: ?(see Supplemental Data with the online version of this article)?.

%Note that if authors wish to submit long DNA sequence appendices as supplemental material, they should select the "DNA sequences (online-only supplemental)" option on Editorial Manager. This ensures that lengthy appendices are not built into the reviewers' PDF, but are still accessible to the reviewers.
\documentclass[11pt]{article}
\usepackage[letterpaper, margin=1in]{geometry} 
\usepackage{hyperref}
\usepackage[leftcaption]{sidecap}
\sidecaptionvpos{figure}{c} %position side caption
\usepackage{lineno}
\usepackage{lipsum}
\usepackage{indentfirst}
\usepackage{pdflscape}
\usepackage{color}
\usepackage{multirow}
\usepackage[utf8]{inputenc}
\usepackage{array}
	\newcounter{rowno}
	\setcounter{rowno}{0}
\usepackage{booktabs}
\usepackage[pdftex]{graphicx}  
\usepackage{caption}
\usepackage{natbib}
\usepackage{authblk}
\usepackage[table]{xcolor} %for table line color alteration
\usepackage{subscript}
\usepackage{longtable}
\usepackage{fancyhdr}
	\renewcommand{\headrulewidth}{0pt}
	\rfoot{\thepage}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt}
\cfoot{\thepage}

\pagestyle{fancy}

\begin{document}

\chead{\small{Vann et al.---American Journal of Botany (\#\#)(\#): \#\#\#-\#\#\#. 2014. -- Appendix S1 -- Page \thepage}}

\centering{Appendix S1: Accessions of \emph{Zea mays} ssp. \emph{mexicana} (RIMME) and \emph{Zea mays} ssp. \emph{parviglumis} (RIMPA) sampled. RIHY is a \emph{Z. mays} ssp. \emph{parviglumis} and \emph{Zea mays} ssp. \emph{mays} hybrid.}

% Appendix S1. Parv and mex samples 
\begin{scriptsize}  % Switch from 12pt to 11pt; otherwise, table won't fit
\rowcolors{1}{gray!25}{white}
\begin{longtable}{>{\stepcounter{rowno}}cccccc}
\hiderowcolors
\hline
\multicolumn{1}{c}{\textbf{Accession}} & \multicolumn{1}{c}{\textbf{USDA ID}} & \multicolumn{1}{c}{\textbf{Locality}} & \multicolumn{1}{c}{\textbf{Alleles Sampled}} & \multicolumn{1}{c}{\textbf{\emph{Hopscotch} Freq.}} & \multicolumn{1}{c}{\textbf{No \emph{Hopscotch} Freq.}} \\
\hline
\endhead
\showrowcolors
    RIHY0009 & N/A   & N/A   & 2     & 0.5   & 0.5 \\
    RIMME0006 & 566673 & Durango & 2	& 0 & 1 \\
    RIMME0007 & 566680 & Guanajuato & 2 & 0 & 1 \\
    RIMME0008 & 566681 & Michoacan & 2 & 0 & 1 \\
    RIMME0009 & 566682 & Distrito Federal & 2 & 0 & 1 \\
    RIMME0011 & 566685 & Mexico & 2     & 0     & 1 \\
    RIMME0014 & 714151 & Breeders line & 6     & 0     & 1 \\
    RIMME0017 & 699874 & Ayotlan & 8     & 0     & 1 \\
    RIMME0021 & N/A   & El Porvenir & 69    & 0.17 & 0.83 \\
    RIMME0026 & N/A   & Opopeo & 42    & 0.07 & 0.93 \\
    RIMME0028 & N/A   & Puruandiro & 28    & 0.04 & 0.96 \\
    RIMME0029 & N/A   & Ixtlan & 35    & 0     & 1 \\
    RIMME0030 & N/A   & San Pedro & 27    & 0     & 1 \\
    RIMME0031 & N/A   & Tenango del Aire & 25    & 0.08  & 0.92 \\
    RIMME0032 & N/A   & Nabogame & 24    & 0     & 1 \\
    RIMME0033 & N/A   & Puerta Encantada & 25    & 0     & 1 \\
    RIMME0034 & N/A   & Santa Clara & 23    & 0     & 1 \\
    RIMME0035 & N/A   & Xochimilco & 25    & 0     & 1 \\
    RIMPA0001 & 87168 & El Salado & 4     & 0     & 1 \\
    RIMPA0003 & 87171 & Mazatlan & 8 & 0.13 & 0.87 \\
    RIMPA0017 & 87200 & N/A   & 4     & 0     & 1 \\
    RIMPA0019 & 87213 & El Salado & 2     & 0.50   & 0.50 \\
    RIMPA0029 & 87244 & N/A   & 2     & 0.50   & 0.50 \\
    RIMPA0031 & 87249 & N/A   & 2     & 0.5   & 0.5 \\
    RIMPA0035 & 87288 & Jalisco & 4     & 0     & 1 \\
    RIMPA0040 & 288185 & Mexico & 4     & 0     & 1 \\
    RIMPA0042 & 288187 & Guerrero & 4     & 0.25  & 0.75 \\
    RIMPA0043 & 288188 & Guerrero & 4     & 0     & 1 \\
    RIMPA0045 & 288193 & Guerrero & 4     & 0     & 1 \\
    RIMPA0055 & 714152 & Breeders line & 2     & 0     & 1 \\
    RIMPA0056 & 714153 & Breeders line & 2     & 0.50   & 0.50 \\
    RIMPA0057 & 714154 & Breeders line & 2     & 0.50   & 0.50 \\
    RIMPA0058 & N/A   & N/A   & 4     & 0.50   & 0.50 \\
    RIMPA0059 & N/A   & N/A   & 4     & 1     & 0 \\
    RIMPA0060 & 714157 & Breeders line & 2     & 0     & 1 \\
    RIMPA0061 & 714158 & Breeders line & 4     & 0.5   & 0.5 \\
    RIMPA0062 & 714159 & Breeders line & 4     & 0.5   & 0.5 \\
    RIMPA0063 & 714160 & Breeders line & 4     & 0     & 1 \\
    RIMPA0064 & 714161 & Breeders line & 3     & 0     & 1 \\
    RIMPA0065 & 714162 & Breeders line & 4     & 0.25  & 0.75 \\
    RIMPA0068 & 699861 & Jalisco, Mexico & 16    & 0     & 1 \\
    RIMPA0069 & 699862 & Ixtlan & 14    & 0.14 & 0.86 \\
    RIMPA0070 & 699863 & Benito Jaurez & 16    & 0     & 1 \\
    RIMPA0071 & 699864 & Tuzantla & 28    & 0     & 1 \\
    RIMPA0072 & 699865 & Tiquicheo & 16    & 0     & 1 \\
    RIMPA0073 & 699866 & Tiquicheo & 16    & 0.12 & 0.88 \\
    RIMPA0074 & 699867 & Huetamo & 12    & 0     & 1 \\
    RIMPA0075 & 699868 & Huetamo & 2     & 0     & 1 \\
    RIMPA0076 & 699869 & Huetamo & 4     & 0     & 1 \\
    RIMPA0077 & 699870 & Caracuaro & 2     & 0     & 1 \\
    RIMPA0078 & 699871 & Caracuaro & 2     & 0.5   & 0.5 \\
    RIMPA0079 & 699872 & Villa Madero & 14    & 0     & 1 \\
    RIMPA0080 & 699873 & Guachinango & 12    & 0     & 1 \\
    RIMPA0081 & 699875 & Ameca & 16    & 0     & 1 \\
    RIMPA0083 & 699877 & Tepoztlan & 14    & 0     & 1 \\
    RIMPA0084 & 699878 & Tepoztlan & 16    & 0     & 1 \\
    RIMPA0085 & 699879 & Miahuatlan & 16    & 0     & 1 \\
    RIMPA0086 & 699880 & Miahuatlan & 16    & 0.06 & 0.94 \\
    RIMPA0087 & 699881 & Tecoanapa & 24    & 0     & 1 \\
    RIMPA0089 & 699883 & Guerrero & 12    & 0     & 1 \\
    RIMPA0090 & 699884 & Guerrero & 10    & 0     & 1 \\
    RIMPA0091 & 699885 & Guerrero & 16    & 0     & 1 \\
    RIMPA0092 & 699886 & Guerrero & 10    & 0     & 1 \\
    RIMPA0093 & 699887 & Guerrero & 26    & 0.08 & 0.92 \\
    RIMPA0094 & 699888 & Guerrero & 2     & 0     & 1 \\
    RIMPA0095 & 699889 & Guerrero & 4     & 0     & 1 \\
    RIMPA0096 & 699890 & Guerrero & 26    & 0.04 & 0.96 \\
    RIMPA0097 & 699891 & Guerrero & 6     & 0     & 1 \\
    RIMPA0098 & 699892 & Guerrero & 4     & 0     & 1 \\
    RIMPA0099 & 699893 & Guerrero & 4     & 0     & 1 \\
    RIMPA0100 & 699894 & Guerrero & 6     & 0     & 1 \\
    RIMPA0101 & 699895 & Guerrero & 2     & 0     & 1 \\
    RIMPA0103 & 699897 & Guerrero & 2     & 0     & 1 \\
    RIMPA0104 & 699898 & Guerrero & 22    & 0.09 & 0.91 \\
    RIMPA0105 & 699899 & Guerrero & 6     & 0     & 1 \\
    RIMPA0106 & 699900 & Guerrero & 6     & 0.33 & 0.67 \\
    RIMPA0107 & 699901 & Guerrero & 4     & 0     & 1 \\
    RIMPA0108 & 699902 & Guerrero & 6     & 0     & 1 \\
    RIMPA0109 & 699903 & Michoacan & 4     & 0.25  & 0.75 \\
    RIMPA0110 & 699904 & Michoacan & 2     & 0     & 1 \\
    RIMPA0111 & 699905 & Michoacan & 4     & 0     & 1 \\
    RIMPA0112 & 699906 & Michoacan & 4     & 0.25  & 0.75 \\
    RIMPA0114 & 699908 & Michoacan & 6     & 0.17 & 0.83 \\
    RIMPA0116 & 699910 & Mexico & 2     & 0     & 1 \\
    RIMPA0117 & 699911 & Mexico & 4     & 0     & 1 \\
    RIMPA0118 & 699912 & Mexico & 6     & 0.17 & 0.83 \\
    RIMPA0119 & 699913 & Mexico & 2     & 0     & 1 \\
    RIMPA0120 & 699914 & Mexico & 1     & 1     & 0 \\
    RIMPA0121 & 699915 & Mexico & 2     & 0     & 1 \\
    RIMPA0128 & 699922 & Mexico & 2     & 0.5   & 0.5 \\
    RIMPA0129 & 699923 & Michoacan & 2     & 0.5   & 0.5 \\
    RIMPA0135 & 699929 & Nayarit & 24    & 0     & 1 \\
    RIMPA0138 & 699932 & Jalisco & 2     & 0.5   & 0.5 \\
    RIMPA0139 & 699933 & Jalisco & 1     & 1     & 0 \\
    RIMPA0142 & 699936 & Colima & 18    & 0.44 & 0.56 \\
    RIMPA0144 & 699938 & Jalisco & 2     & 1     & 0 \\
    RIMPA0145 & 699939 & Michoacan & 1     & 1     & 0 \\
    RIMPA0147 & 699941 & Jalisco & 1     & 1     & 0 \\
    RIMPA0155 & N/A   & Jalisco & 73    & 0.01 & 0.99 \\
    RIMPA0156 & N/A   & Jalisco & 20    & 0     & 1 \\
    RIMPA0157 & N/A   & Jalisco & 58    & 0.34 & 0.66 \\
    RIMPA0158 & N/A   & Jalisco & 64    & 0.53 & 0.47 \\
    RIMPA0159 & N/A   & Jalisco & 26    & 0     & 1 \\
    RIMPA0162 & 21785 & N/A   & 4     & 0     & 1 \\
    \hline
\end{longtable}
\end{scriptsize}
\clearpage

\pagenumbering{arabic}
\chead{\small{Vann et al.---American Journal of Botany (\#\#)(\#): \#\#\#-\#\#\#. 2014. -- Appendix S2 -- Page \thepage}}
% Appendix S2. Mays sampling

\centering{Appendix S2: \emph{Hopscotch} frequency in sampled \emph{Zea mays} ssp. \emph{mays} (RIMMA).}

\begin{scriptsize}  % Switch from 12pt to 11pt; otherwise, table won't fit
\rowcolors{1}{gray!25}{white}
\begin{longtable}{>{\stepcounter{rowno}}ccc}
\hiderowcolors
\hline
\multicolumn{1}{c}{\textbf{Accession}} & \multicolumn{1}{c}{\textbf{Alleles Sampled}} & \multicolumn{1}{c}{\textbf{\emph{Hopscotch} Freq.}} \\
\hline
\endhead
\showrowcolors
    RIMMA0066 & 2     & 1 \\
    RIMMA0075 & 2     & 1 \\
    RIMMA0077 & 2     & 1 \\
    RIMMA0079 & 2     & 1 \\
    RIMMA0081 & 2     & 1 \\
    RIMMA0084 & 2     & 1 \\
    RIMMA0086 & 2     & 1 \\
    RIMMA0088 & 2     & 1 \\
    RIMMA0089 & 2     & 1 \\
    RIMMA0090 & 2     & 1 \\
    RIMMA0092 & 4     & 1 \\
    RIMMA0094 & 4     & 1 \\
    RIMMA0097 & 2     & 1 \\
    RIMMA0099 & 2     & 1 \\
    RIMMA0100 & 2     & 1 \\
    RIMMA0101 & 2     & 1 \\
    RIMMA0104 & 2     & 1 \\
    RIMMA0108 & 2     & 1 \\
    RIMMA0111 & 6     & 1 \\
    RIMMA0115 & 2     & 1 \\
    RIMMA0117 & 2     & 1 \\
    RIMMA0130 & 2     & 1 \\
    RIMMA0133 & 2     & 1 \\
    RIMMA0134 & 2     & 1 \\
    RIMMA0135 & 2     & 1 \\
    RIMMA0142 & 2     & 0.5 \\
    RIMMA0143 & 4     & 1 \\
    RIMMA0146 & 4     & 1 \\
    RIMMA0149 & 2     & 1 \\
    RIMMA0152 & 2     & 1 \\
    RIMMA0153 & 2     & 1 \\
    RIMMA0154 & 2     & 1 \\
    RIMMA0155 & 2     & 1 \\
    RIMMA0156 & 2     & 1 \\
    RIMMA0157 & 2     & 1 \\
    RIMMA0158 & 2     & 1 \\
    RIMMA0159 & 2     & 1 \\
    RIMMA0160 & 2     & 1 \\
    RIMMA0162 & 2     & 1 \\
    RIMMA0166 & 2     & 1 \\
    RIMMA0167 & 2     & 1 \\
    RIMMA0168 & 2     & 1 \\
    RIMMA0169 & 2     & 1 \\
    RIMMA0172 & 2     & 1 \\
    RIMMA0174 & 4     & 1 \\
    RIMMA0177 & 2     & 1 \\
    RIMMA0178 & 2     & 1 \\
    RIMMA0179 & 2     & 1 \\
    RIMMA0181 & 2     & 1 \\
    RIMMA0183 & 2     & 1 \\
    RIMMA0184 & 2     & 1 \\
    RIMMA0186 & 2     & 1 \\
    RIMMA0187 & 1     & 1 \\
    RIMMA0188 & 2     & 1 \\
    RIMMA0195 & 2     & 1 \\
    RIMMA0196 & 2     & 1 \\
    RIMMA0197 & 2     & 1 \\
    RIMMA0198 & 2     & 1 \\
    RIMMA0199 & 2     & 1 \\
    RIMMA0200 & 2     & 1 \\
    RIMMA0202 & 2     & 1 \\
    RIMMA0203 & 2     & 1 \\
    RIMMA0206 & 2     & 1 \\
    RIMMA0208 & 2     & 1 \\
    RIMMA0209 & 2     & 1 \\
    RIMMA0210 & 2     & 1 \\
    RIMMA0212 & 2     & 1 \\
    RIMMA0213 & 2     & 1 \\
    RIMMA0214 & 2     & 1 \\
    RIMMA0217 & 2     & 1 \\
    RIMMA0218 & 2     & 1 \\
    RIMMA0220 & 2     & 1 \\
    RIMMA0221 & 2     & 1 \\
    RIMMA0222 & 2     & 1 \\
    RIMMA0223 & 2     & 1 \\
    RIMMA0226 & 2     & 1 \\
    RIMMA0227 & 2     & 1 \\
    RIMMA0228 & 2     & 1 \\
    RIMMA0229 & 2     & 1 \\
    RIMMA0230 & 2     & 1 \\
    RIMMA0232 & 2     & 1 \\
    RIMMA0233 & 2     & 1 \\
    RIMMA0235 & 2     & 0.5 \\
    RIMMA0242 & 2     & 1 \\
    RIMMA0243 & 2     & 1 \\
    RIMMA0247 & 4     & 1 \\
    RIMMA0248 & 2     & 1 \\
    RIMMA0249 & 2     & 1 \\
    RIMMA0252 & 2     & 1 \\
    RIMMA0253 & 2     & 1 \\
    RIMMA0254 & 2     & 1 \\
    RIMMA0256 & 2     & 1 \\
    RIMMA0257 & 2     & 1 \\
    RIMMA0258 & 2     & 1 \\
    RIMMA0259 & 2     & 1 \\
    RIMMA0260 & 2     & 1 \\
    RIMMA0262 & 2     & 1 \\
    RIMMA0263 & 2     & 1 \\
    RIMMA0264 & 2     & 1 \\
    RIMMA0265 & 2     & 1 \\
    RIMMA0268 & 2     & 1 \\
    RIMMA0269 & 2     & 1 \\
    RIMMA0270 & 2     & 1 \\
    RIMMA0272 & 2     & 1 \\
    RIMMA0275 & 2     & 1 \\
    RIMMA0276 & 2     & 1 \\
    RIMMA0277 & 2     & 1 \\
    RIMMA0279 & 2     & 1 \\
    RIMMA0280 & 2     & 1 \\
    RIMMA0283 & 2     & 1 \\
    RIMMA0285 & 2     & 1 \\
    RIMMA0288 & 2     & 1 \\
    RIMMA0290 & 2     & 1 \\
    RIMMA0291 & 2     & 1 \\
    RIMMA0292 & 2     & 1 \\
    RIMMA0293 & 2     & 1 \\
    RIMMA0298 & 2     & 1 \\
    RIMMA0302 & 2     & 1 \\
    RIMMA0305 & 2     & 1 \\
    RIMMA0310 & 2     & 1 \\
    RIMMA0312 & 2     & 1 \\
    RIMMA0320 & 2     & 1 \\
    RIMMA0322 & 2     & 1 \\
    RIMMA0324 & 2     & 1 \\
    RIMMA0334 & 2     & 1 \\
    RIMMA0336 & 2     & 1 \\
    RIMMA0337 & 2     & 1 \\
    RIMMA0338 & 2     & 1 \\
    RIMMA0339 & 2     & 1 \\
    RIMMA0340 & 2     & 1 \\
    RIMMA0341 & 2     & 1 \\
    RIMMA0342 & 2     & 1 \\
    RIMMA0344 & 2     & 1 \\
    RIMMA0346 & 2     & 1 \\
    RIMMA0347 & 2     & 1 \\
    RIMMA0350 & 2     & 1 \\
    RIMMA0351 & 2     & 1 \\
    RIMMA0352 & 2     & 1 \\
    RIMMA0357 & 2     & 1 \\
    RIMMA0360 & 18    & 1 \\
    RIMMA0362 & 8     & 1 \\
    RIMMA0363 & 16    & 1 \\
    RIMMA0370 & 17    & 1 \\
    RIMMA0372 & 24    & 1 \\
    RIMMA0373 & 22    & 1 \\
    RIMMA0408 & 2     & 1 \\
    RIMMA0411 & 1     & 1 \\
    RIMMA0413 & 1     & 1 \\
    RIMMA0414 & 1     & 1 \\
    RIMMA0424 & 8     & 0.9 \\
    RIMMA0432 & 1     & 1 \\
    RIMMA0434 & 2     & 0.5 \\
    RIMMA0435 & 1     & 1 \\
    RIMMA0442 & 2     & 1 \\
    RIMMA0443 & 2     & 1 \\
    RIMMA0444 & 2     & 1 \\
    RIMMA0445 & 2     & 1 \\
    RIMMA0446 & 2     & 1 \\
    RIMMA0447 & 2     & 1 \\
    RIMMA0448 & 2     & 1 \\
    RIMMA0449 & 2     & 1 \\
    RIMMA0450 & 2     & 1 \\
    RIMMA0451 & 2     & 1 \\
    RIMMA0452 & 2     & 1 \\
    RIMMA0453 & 2     & 1 \\
    RIMMA0454 & 2     & 1 \\
    RIMMA0455 & 2     & 1 \\
    RIMMA0456 & 2     & 1 \\
    RIMMA0457 & 2     & 1 \\
    RIMMA0458 & 2     & 1 \\
    RIMMA0459 & 2     & 1 \\
    RIMMA0483 & 2     & 1 \\
    RIMMA0490 & 2     & 1 \\
    RIMMA0515 & 2     & 1 \\
    RIMMA0537 & 2     & 1 \\
    RIMMA0550 & 2     & 1 \\
    RIMMA0553 & 2     & 1 \\
    RIMMA0559 & 2     & 1 \\
    RIMMA0561 & 2     & 1 \\
    RIMMA0562 & 2     & 1 \\
    RIMMA0571 & 2     & 1 \\
    RIMMA0572 & 2     & 1 \\
    RIMMA0577 & 2     & 1 \\
    RIMMA0579 & 2     & 1 \\
    RIMMA0590 & 2     & 1 \\
    RIMMA0591 & 2     & 1 \\
    RIMMA0592 & 2     & 1 \\
    RIMMA0593 & 2     & 1 \\
    RIMMA0594 & 2     & 1 \\
    RIMMA0595 & 2     & 1 \\
    RIMMA0596 & 2     & 1 \\
    RIMMA0597 & 2     & 1 \\
    RIMMA0598 & 2     & 1 \\
    RIMMA0599 & 2     & 1 \\
    RIMMA0600 & 2     & 1 \\
    RIMMA0601 & 2     & 1 \\
    RIMMA0602 & 2     & 1 \\
    RIMMA0603 & 2     & 1 \\
    RIMMA0604 & 2     & 1 \\
    RIMMA0605 & 2     & 1 \\
    RIMMA0606 & 2     & 1 \\
    RIMMA0607 & 2     & 1 \\
    RIMMA0608 & 2     & 1 \\
    RIMMA0609 & 2     & 1 \\
    RIMMA0610 & 2     & 1 \\
    RIMMA0611 & 2     & 1 \\
    RIMMA0612 & 2     & 1 \\
    RIMMA0613 & 2     & 1 \\
    RIMMA0622 & 2     & 0.5 \\
    RIMMA0624 & 1     & 1 \\
    RIMMA0629 & 2     & 0.5 \\
    RIMMA0631 & 2     & 0.5 \\
    RIMMA0659 & 2     & 1 \\
    RIMMA0660 & 2     & 1 \\
    RIMMA0678 & 1     & 1 \\
    RIMMA0681 & 2     & 1 \\
    RIMMA0683 & 2     & 1 \\
    RIMMA0684 & 2     & 1 \\
    RIMMA0685 & 4     & 1 \\
    RIMMA0693 & 2     & 1 \\
    RIMMA0694 & 2     & 1 \\
    RIMMA0695 & 1     & 1 \\
    RIMMA0699 & 2     & 1 \\
    RIMMA0704 & 2     & 1 \\
    RIMMA0706 & 2     & 1 \\
    RIMMA0711 & 1     & 1 \\
    RIMMA0713 & 2     & 1 \\
    RIMMA0715 & 2     & 1 \\
    RIMMA0717 & 2     & 1 \\
    RIMMA0718 & 2     & 1 \\
    RIMMA0723 & 2     & 1 \\
    RIMMA0724 & 2     & 1 \\
    RIMMA0725 & 1     & 1 \\
    RIMMA0728 & 1     & 1 \\
    RIMMA0732 & 2     & 1 \\
    RIMMA0734 & 2     & 1 \\
    RIMMA0738 & 2     & 1 \\
    RIMMA0739 & 2     & 1 \\
    RIMMA0744 & 2     & 1 \\
    RIMMA0747 & 1     & 1 \\
    RIMMA0748 & 2     & 1 \\
    RIMMA0749 & 2     & 1 \\
    RIMMA0750 & 2     & 0.5 \\
    RIMMA0751 & 1     & 1 \\
    RIMMA0752 & 1     & 1 \\
    RIMMA0753 & 2     & 1 \\
    RIMMA0755 & 2     & 1 \\
    \hline
\end{longtable}
\end{scriptsize}

\clearpage
\pagenumbering{arabic}
\chead{\small{Vann et al.---American Journal of Botany (\#\#)(\#): \#\#\#-\#\#\#. 2014. -- Appendix S3 -- Page \thepage}}
% Appendix S3. Cartoon representation of tb1 locus
%-------------------------------------------------------------------
\begin{figure*}[!t]
  \begin{center}
   \includegraphics[width=180mm]{FigS1LocusCartoon.pdf}
   \flushleft{Appendix S3: Representation of the upstream regulatory region of \emph{tb1}, showing the \emph{tb1} coding region (green) and the \emph{Hopscotch} insertion (red). Arrows show the location of primer sets; in black, primers used for amplification and sequencing (Region 1; within the 5' UTR, and Region 2; 66,169 bp upstream from the tb1 ORF); in blue, primers used to genotype the \emph{Hopscotch} insertion.}
  \end{center}
\end{figure*}
%-------------------------------------------------------------------

\clearpage
\pagenumbering{arabic}
\chead{\small{Vann et al.---American Journal of Botany (\#\#)(\#): \#\#\#-\#\#\#. 2014. -- Appendix S4 -- Page \thepage}}

% Appendix S4. Gel image
%-------------------------------------------------------------------
\begin{figure*}[!t]
  \begin{center}
   \includegraphics[width=100mm]{FigS2Gel.png}
   \flushleft{Appendix S4: Agarose gel image of amplification products using our primer sets. Genotypes are indicated at the top of the gel.} 
  \end{center}
\end{figure*}
%-------------------------------------------------------------------

\clearpage
\pagenumbering{arabic}
\chead{\small{Vann et al.---American Journal of Botany (\#\#)(\#): \#\#\#-\#\#\#. 2014. -- Appendix S5 -- Page \thepage}}

% Appendix S5. NJ Trees
%-------------------------------------------------------------------
\begin{figure*}[!t]
  \begin{center}
   \includegraphics[width=150mm]{FigS3NJtrees.pdf}
    \flushleft{Appendix S5: Neighbor-joining tree of the sequenced region in the 5' UTR (right; Region 1) and the 66,169 bp upstream region (left; Region 2) of \emph{tb1}. Individuals with genotype data are colored: Homozygous for the teosinte (no \emph{Hopscotch}) allele (red), homozygous for the maize (\emph{Hopscotch}) allele (blue), heterozygotes (purple). TILs (teosinte inbred lines) are colored in green, with stars indicating the 3 TILs known to have the \emph{Hopscotch} insertion. Black indicates individuals not genotyped for the \emph{Hopscotch} insertion.} 
  \end{center}
\end{figure*}
%-------------------------------------------------------------------
\end{document}